---
title: "Intro to Tidyverse WS-1"
author: "Amanda"
date: "2023-10-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install packages and load libraries
```{r}
#install.packages("tidyverse")
#install.packages("skimr")

library(tidyverse)
library(skimr)
```

Read the dataset into our environment.

Step 1. Download it from this website, and then upload it into Rstudio cloud - 
 https://www.kaggle.com/datasets/gabrielsantello/wholesale-and-retail-orders-dataset/code

Step 2. Assign the csv files to the dataframes 
```{r}
orders <- read_csv("orders.csv")
product <- read_csv("product-supplier.csv")

```

Look at the overall structure of the datasets
```{r}
glimpse(orders)
glimpse(product)

```

use the head function to see the first few rows or the tail function to see the last few rows
```{r}
head(orders, 10)
tail(product, 5)

```

We will use a function from the janitor package to clean up the variable names.
Remember - install.packages once and activate the libraries every script/rmd file

```{r}
#install.packages("janitor")
library(janitor)
```

Clean the variable names in the orders dataframe
```{r}
orders <- clean_names(orders)
glimpse(orders)

```

Convert the date columns to from character types to date format. We will use the Lubridate package to do this - this has already been installed and activated as a part of the tidyverse.

```{r}
orders$delivery_date <- dmy(orders$delivery_date)
orders$date_order_was_placed <- dmy(orders$date_order_was_placed)
glimpse(orders)

```

Plot a histogram of the order values
```{r}
ggplot(data = orders, aes(x = total_retail_price_for_this_order)) + geom_histogram()

```

This is quite skewed with a few large value orders. Lets look at this a bit closer

```{r}

orders %>% filter(total_retail_price_for_this_order <=2000) %>% 
  ggplot(aes(x = total_retail_price_for_this_order))+
  geom_histogram(bins = 100) + theme_light()

```

Review Customer status
```{r}
orders %>% count(customer_status)
```

Clean the values and convert them to factors

```{r}
orders$customer_status <- str_to_lower(orders$customer_status)
orders %>% count(customer_status)

orders$customer_status <- as_factor(orders$customer_status)

glimpse(orders)

```

How to apply this across multiple columns - but you will need to assign it to the data frame to apply

```{r}
orders <- orders %>%  
  mutate(across(all_of(c("customer_id","product_id")), as.factor))
glimpse(orders)

```

How many orders and products per order
```{r}
orders %>% count(order_id, sort = TRUE)

orders %>% count(product_id, sort = TRUE)

orders %>% count(order_id, product_id, sort = TRUE)
```

Find the retail value of the items and add this as a new column
```{r}

orders <- orders %>% mutate(product_retail_value = total_retail_price_for_this_order/quantity_ordered)

average_orders <- orders %>% 
  group_by(product_id) %>% 
  mutate(cost_price_per_unit = mean(cost_price_per_unit)) %>% 
  mutate(product_retail_value = mean(product_retail_value)) %>% 
  arrange(desc(product_retail_value)) %>% 
  distinct(product_id, cost_price_per_unit, product_retail_value)
```

Lets join the two dataframes together on the product_id column. First clean the names and convert the product_id in the product df to a factor.
```{r}
glimpse(product)

product <- clean_names(product)

product <- product %>%  
  mutate(across(all_of(c("product_id","supplier_id")), as.factor))

glimpse(product)

sales <- inner_join(orders, product)

```

Calculate and plot the profit margin
```{r}
sales <- sales %>% 
  mutate(sales_margin = product_retail_value - cost_price_per_unit)

sales <- sales %>% rename(wholesale_price = cost_price_per_unit, retail_price = product_retail_value)
glimpse(sales)

library(viridis)
sales %>% ggplot(aes(x = wholesale_price, y = retail_price, colour = product_category))+ geom_point()+ theme_minimal() + scale_colour_viridis(discrete = TRUE, option = "D")+
  facet_wrap(~ product_category, ncol = 3)
```












